Do Alternate Wetting and Drying Irrigation Technology and Nitrogen Rates Affect Rice Sheath Blight?

### 2018-05-02 analysis/paper/paper.Rmd
More edits to statistical analysis section
New para for describing the model runs and checking
Edits to Discussion section

Area under the disease progress stairs (AUDPS) [@simko2012area] was calculated for both disease incidence and severity using the R package `agricolae` [@R2018; @deMendiburu2017]. Because the severity data were collected on an unevenly spaced categorical scale, they were converted to the midpoint value of the percent range.

As most of the data's residuals did not meet assumptions for normality, the analysis was carried out using Bayesian multivariate generalised linear mixed models implemented in the R package `MCMCglmm` [@MCMCglmm2010]. Six models were created where the the dependent variables were 2015 tiller ShB incidence, 2015 tiller ShB severity, 2015 leaf ShB severity; 2016 tiller ShB incidence, 2016 tiller ShB severity; 2016 leaf ShB severity, respectively. The independent variables were the irrigation management and N rate with replicate treated as a random variable. The base levels (control treatments) for the analyses for 2015 were: N rate, 0 kg/ha; irrigation management, FLD and for 2016 were: N rate, 60 kg/ha; irrigation management, FLD. Four MCMC chains were run for 55,000 steps with the first 5000 discarded as burn in. The remaining 50,000 samples were thinned by taking every tenth sample resulting in 5,000 independent draws from the posterior distribution of the parameters of the model. The `coda` [@coda2006] package was used to provide utilities to check convergence using visual inspection of graphs and the Gelman-Rubin test. All data generated from this project and code used for analysis are available as a reproducible research compendium from <https://github.com/openplantpathology/rice_awd_pests> [@Sparks2018].

@@

In both experiments we were unable to detect any significant effect of AWD on ShB that would cause increases in the disease, which could hinder adoption of the technology. In only one instance were we able to detect any effect of AWD on ShB (Fig. \@ref(fig:estimates-plot)d), where in the 2016 experiment the tiller ShB severity was significantly lower for AWD than the traditional FLD treatment. This may may indicate a possible adverse effect of using AWD on tiller ShB severity.

The effects of N rates on sheath blight were to be expected, where higher rates of N caused an increase in disease incidence and severity in 2015 and in tiller incidence in 2016. However, in 2016 tiller and leaf severity had no detectable differences due to the N rate.

>>>>

Area under the disease progress stairs (AUDPS) [@simko2012area] was calculated for both disease incidence and severity using the R package `agricolae` [@R2018; @deMendiburu2017]. The severity data were converted to the midpoint perctn value of the scale range.

As most of the data's residuals did not meet assumptions for normality, the analysis was carried out using Bayesian multivariate generalised linear mixed models implemented in the R package `MCMCglmm` [@MCMCglmm2010]. Six models were created where the the dependent variables were 2015 tiller ShB incidence, 2015 tiller ShB severity, 2015 leaf ShB severity; 2016 tiller ShB incidence, 2016 tiller ShB severity; 2016 leaf ShB severity, respectively. The independent variables were the irrigation management and N rate with replicate treated as a random variable. The base levels (control treatments) for the analyses for 2015 were: N rate, 0 kg/ha; irrigation management, FLD and for 2016 were: N rate, 60 kg/ha; irrigation management, FLD.

Four MCMC chains were run for 55,000 steps with the first 5000 discarded as burn in. The remaining 50,000 samples were thinned by taking every tenth sample resulting in 5,000 independent draws from the posterior distribution of the parameters of the model. The `coda` [@coda2006] package was used to provide utilities to check convergence using visual inspection of graphs and the Gelman-Rubin test. All data generated from this project and code used for analysis are available as a reproducible research compendium from <https://github.com/openplantpathology/rice_awd_pests> [@Sparks2018].

@@

In both experiments we were unable to detect any significant effect of AWD on ShB that would cause increases in the disease, which could hinder adoption of the technology. In only one instance were we able to detect any effect of AWD on ShB. In the 2016 experiment the tiller ShB severity was significantly lower for AWD than FLD (Fig. \@ref(fig:ShB-W-box)d, \@ref(fig:estimates-plot)d), indicating a possible adverse effect of using AWD on tiller ShB severity under conditions where there is high sheath blight pressure.

The findings of the effects of N rates on sheath blight were to be expected, where higher rates of N caused an increase in disease incidence and severity in 2015 and in tiller incidence in 2016. However, in 2016 tiller and leaf severity had no detectable differences due to the N rate. However, it should be noted that the levels of leaf severity remained low (< 1%) throughout the growing season in all treatments for both years.
